Supplementary Information

### 2020-11-15 doc/manuscript/MEE_manuscript_SI.tex
organizing appendices

\hypertarget{appendix-s2.-methods-for-comparing-climwin-results-with-traditional-methods}{%
\subsection{Appendix S2. Methods for comparing climwin results with
traditional
methods}\label{appendix-s2.-methods-for-comparing-climwin-results-with-traditional-methods}}

@@

\hypertarget{appendix-s3.-dealing-with-rapidly-changing-climate-and-tree-growth}{%
\subsection{Appendix S3. Dealing with rapidly changing climate and tree
growth}\label{appendix-s3.-dealing-with-rapidly-changing-climate-and-tree-growth}}

@@

ABAL & Pinaceae & Abies alba & ZOF & needleleaf & evergreen & shade-tolerant & neglected in Zofin\\

@@

FASY & Fagaceae & Fagus sylvatica & ZOF & broadleaf & deciduous (cold) & shade-tolerant & neglected in Zofin\\

@@

\begin{tabular}{llrrrrlll}
\textbf{site} & \textbf{species code} & \textbf{n trees all} & \textbf{n cores all} & \textbf{n trees dbh} & \textbf{n cores dbh} & \textbf{dbh range sampled} & \textbf{dbh range reconstructed*} & \textbf{date range}\\
BCI & JACO & 12 & 18 & 11 & 17 & 30.2-63.5 & 2.6-56.4 & 1931-2014\\
BCI & TEPA & 18 & 29 & 17 & 26 & 22.1-59.5 & 2.7-49.4 & 1931-2014\\
BCI & TRTU & 23 & 37 & 20 & 31 & 20.7-43.6 & 4.8-41.5 & 1931-2014\\
CB & ABBI & 22 & 41 & 20 & 37 & 13.9-54.2 & 0.9-46.4 & 1903-2000\\
CB & PIEN & 12 & 21 & 10 & 15 & 14-54.9 & 0.9-33.1 & 1903-2000\\
CB & PIFL & 13 & 22 & 12 & 21 & 17.6-64.1 & 1.5-47.5 & 1903-1998\\
CB & PILO & 11 & 16 & 8 & 12 & 45.8-121.6 & 35.4-118.5 & 1903-1999\\
CB & PIPU & 15 & 28 & 15 & 28 & 22.4-50.8 & 9.5-48.4 & 1903-2000\\
CB & POTR & 17 & 27 & 17 & 26 & 23.6-47.6 & 4.3-35.4 & 1903-2000\\
CB & PSME & 11 & 20 & 10 & 18 & 20.7-64.2 & 0.5-41.5 & 1903-1999\\
HF & ACRU & 18 & 59 & 18 & 59 & 10.1-22.1 & 0.9-20.4 & 1903-2013\\
HF & BEAL & 13 & 44 & 13 & 44 & 10.2-37.9 & 0-17.2 & 1904-2013\\
HF & QURU & 74 & 180 & 73 & 177 & 19.5-53 & 1.1-48.3 & 1903-2014\\
HF & TSCA & 32 & 83 & 32 & 71 & 10.6-37 & 0-28.4 & 1923-2014\\
HKK & AFXY & 39 & 127 & 39 & 127 & 20.1-98.7 & 0.1-81.4 & 1903-2011\\
HKK & CHTA & 28 & 70 & 28 & 70 & 16-64.6 & 0.2-59.5 & 1904-2010\\
HKK & MEAZ & 46 & 130 & 46 & 130 & 25.6-98.1 & 3.8-80.3 & 1914-2011\\
HKK & TOCI & 45 & 143 & 45 & 143 & 16.6-116.4 & 1.7-80.5 & 1903-2011\\
LDW & ACSA & 35 & 66 & 34 & 64 & 9-64.6 & 0-52.4 & 1903-2019\\
LDW & CAOV & 9 & 18 & 8 & 15 & NA-NA & 1.4-37.4 & 1903-2013\\
LDW & LITU & 15 & 28 & 14 & 25 & NA-NA & 1.2-69.4 & 1903-2019\\
LDW & QUAL & 10 & 20 & NA & NA & NA & NA & 1903-2013\\
LDW & QUMO & 10 & 20 & 8 & 16 & NA-NA & 1.1-52.4 & 1903-2013\\
LDW & QUVE & 9 & 18 & NA & NA & NA & NA & 1903-2013\\
NE & BEPA & 28 & 84 & 28 & 84 & NA-NA & 0-25.5 & 1903-1995\\
NE & JUVI & 29 & 60 & 29 & 60 & 16.6-21.5 & 0.3-18.4 & 1938-1994\\
NE & PIPO & 71 & 134 & 70 & 129 & 19.3-64 & 0-41.1 & 1934-1994\\
LT & PIPO & 10 & 20 & 10 & 20 & 23.2-52.8 & 14.4-48.4 & 1903-2018\\
LT & PIST2 & 7 & 14 & 7 & 14 & 25.7-39.8 & 3.9-34.5 & 1903-2018\\
SCBI & CACO & 15 & 15 & 14 & 14 & 10.62-38.52 & 1.7-32.2 & 1903-2015\\
SCBI & CAGL & 39 & 39 & 36 & 36 & 10.28-52.31 & 1.6-49.3 & 1903-2015\\
SCBI & CAOVL & 25 & 25 & 24 & 24 & 15.11-60.32 & 2.6-47.2 & 1903-2015\\
SCBI & CATO & 15 & 15 & 14 & 14 & 12.86-35.95 & 3.7-28.4 & 1903-2015\\
SCBI & FAGR & 76 & 76 & 74 & 74 & 10.05-41.02 & 0.1-41.2 & 1920-2009\\
SCBI & FRAM & 66 & 66 & 61 & 61 & 8.11-94.73 & 0.1-84.4 & 1903-2016\\
SCBI & FRNI & 12 & 12 & 12 & 12 & 11.04-39.2 & 0.5-27.3 & 1903-1996\\
SCBI & JUNI & 30 & 30 & 28 & 28 & 20.4-76.19 & 5.6-59.5 & 1903-2010\\
SCBI & LITU & 106 & 106 & 104 & 104 & 10-91.42 & 0.1-81.1 & 1903-2010\\
SCBI & PIST & 36 & 36 & 36 & 36 & 13.92-50.96 & 1.6-45.2 & 1931-2010\\
SCBI & QUAL & 66 & 66 & 66 & 66 & 11.4-76.73 & 0.3-70.4 & 1903-2009\\
SCBI & QUMO & 67 & 67 & 67 & 67 & 10.22-84.59 & 0.3-69.5 & 1903-2017\\
SCBI & QURU & 70 & 70 & 70 & 70 & 11.07-87.65 & 2.5-79.2 & 1903-2016\\
SCBI & QUVE & 81 & 81 & 81 & 81 & 16.02-82.33 & 0.5-78.4 & 1903-2009\\
SC & PIMA & 442 & 442 & 101 & 101 & 7-14.9 & 0.5-12.5 & 1903-2013\\
ZOF & ABAL & 55 & 55 & 50 & 50 & 41-121 & 21.3-108.5 & 1903-2011\\
ZOF & FASY & 1369 & 1369 & 1358 & 1358 & NA-NA & 0.1-115.3 & 1903-2013\\
ZOF & PCAB & 644 & 644 & 599 & 599 & NA-NA & 0.5-126.4 & 1903-2011\\

>>>>

\hypertarget{appendix-s2.-methods-for-climate-data-evaluation-and-correction}{%
\subsection{Appendix S2. Methods for climate data evaluation and
correction}\label{appendix-s2.-methods-for-climate-data-evaluation-and-correction}}

\hypertarget{appendix-s3.-methods-for-comparing-climwin-results-with-traditional-methods}{%
\subsection{Appendix S3. Methods for comparing climwin results with
traditional
methods}\label{appendix-s3.-methods-for-comparing-climwin-results-with-traditional-methods}}

@@

\hypertarget{appendix-s4.-dealing-with-rapidly-changing-climate-and-tree-growth}{%
\subsection{Appendix S4. Dealing with rapidly changing climate and tree
growth}\label{appendix-s4.-dealing-with-rapidly-changing-climate-and-tree-growth}}

@@

@@

@@

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{llrrrrlll}
\toprule
\textbf{site} & \textbf{species code} & \textbf{n trees all} & \textbf{n cores all} & \textbf{n trees dbh} & \textbf{n cores dbh} & \textbf{dbh range sampled} & \textbf{dbh range reconstructed*} & \textbf{date range}\\
\midrule
BCI & JACO & 12 & 18 & 11 & 17 & 30.2-63.5 & 2.6-56.4 & 1931-2014\\
\addlinespace
BCI & TEPA & 18 & 29 & 17 & 26 & 22.1-59.5 & 2.7-49.4 & 1931-2014\\
\addlinespace
BCI & TRTU & 23 & 37 & 20 & 31 & 20.7-43.6 & 4.8-41.5 & 1931-2014\\
\addlinespace
CB & ABBI & 22 & 41 & 20 & 37 & 13.9-54.2 & 0.9-46.4 & 1903-2000\\
\addlinespace
CB & PIEN & 12 & 21 & 10 & 15 & 14-54.9 & 0.9-33.1 & 1903-2000\\
\addlinespace
CB & PIFL & 13 & 21 & 12 & 20 & 17.6-64.1 & 1.5-47.5 & 1903-1998\\
\addlinespace
CB & PILO & 17 & 24 & NA & NA & NA & NA & 1903-1999\\
\addlinespace
CB & PIPU & 15 & 28 & 15 & 28 & 22.4-50.8 & 9.5-48.4 & 1903-2000\\
\addlinespace
CB & POTR & 17 & 27 & 17 & 26 & 23.6-47.6 & 4.3-35.4 & 1903-2000\\
\addlinespace
CB & PSME & 11 & 20 & 10 & 18 & 20.7-64.2 & 0.5-41.5 & 1903-1999\\
\addlinespace
HF & ACRU & 18 & 59 & 18 & 59 & 10.1-22.1 & 0.9-20.4 & 1903-2013\\
\addlinespace
HF & BEAL & 13 & 44 & 13 & 44 & 10.2-37.9 & 0-17.2 & 1904-2013\\
\addlinespace
HF & QURU & 74 & 180 & 73 & 177 & 19.5-53 & 1.1-48.3 & 1903-2014\\
\addlinespace
HF & TSCA & 32 & 83 & 32 & 71 & 10.6-37 & 0-28.4 & 1923-2014\\
\addlinespace
HKK & AFXY & 39 & 127 & 39 & 127 & 20.1-98.7 & 0.1-81.4 & 1903-2011\\
\addlinespace
HKK & CHTA & 28 & 70 & 28 & 70 & 16-64.6 & 0.2-59.5 & 1904-2010\\
\addlinespace
HKK & MEAZ & 46 & 130 & 46 & 130 & 25.6-98.1 & 3.8-80.3 & 1914-2011\\
\addlinespace
HKK & TOCI & 45 & 143 & 45 & 143 & 16.6-116.4 & 1.7-80.5 & 1903-2011\\
\addlinespace
LDW & ACSA & 35 & 66 & 34 & 64 & 9-64.6 & 0-52.4 & 1903-2019\\
\addlinespace
LDW & CAOV & 9 & 18 & 8 & 15 & NA-NA & 1.4-37.4 & 1903-2013\\
\addlinespace
LDW & LITU & 15 & 28 & 14 & 25 & NA-NA & 1.2-69.4 & 1903-2019\\
\addlinespace
LDW & QUAL & 10 & 20 & NA & NA & NA & NA & 1903-2013\\
\addlinespace
LDW & QUMO & 10 & 20 & 8 & 16 & NA-NA & 1.1-52.4 & 1903-2013\\
\addlinespace
LDW & QUVE & 9 & 18 & NA & NA & NA & NA & 1903-2013\\
\addlinespace
NE & BEPA & 28 & 84 & 28 & 84 & NA-NA & 0-25.5 & 1903-1995\\
\addlinespace
NE & JUVI & 29 & 60 & 29 & 60 & 16.6-21.5 & 1.6-18.4 & 1941-1994\\
\addlinespace
NE & PIPO & 59 & 114 & 57 & 110 & 19.3-63 & 0-39.5 & 1934-1994\\
\addlinespace
LT & PIPO & 10 & 20 & 10 & 20 & 23.2-52.8 & 14.6-48.4 & 1903-2018\\
\addlinespace
LT & PIST2 & 7 & 14 & 7 & 13 & 25.7-39.8 & 4.2-34.4 & 1903-2018\\
\addlinespace
SCBI & CACO & 15 & 15 & 14 & 14 & 10.62-38.52 & 1.7-32.2 & 1903-2015\\
\addlinespace
SCBI & CAGL & 39 & 39 & 36 & 36 & 10.28-52.31 & 1.6-49.3 & 1903-2015\\
\addlinespace
SCBI & CAOVL & 25 & 25 & 24 & 24 & 15.11-60.32 & 2.6-47.2 & 1903-2015\\
\addlinespace
SCBI & CATO & 15 & 15 & 14 & 14 & 12.86-35.95 & 3.7-28.4 & 1903-2015\\
\addlinespace
SCBI & FAGR & 76 & 76 & 74 & 74 & 10.05-41.02 & 0.1-41.2 & 1920-2009\\
\addlinespace
SCBI & FRAM & 66 & 66 & 61 & 61 & 8.11-94.73 & 0.1-84.4 & 1903-2016\\
\addlinespace
SCBI & FRNI & 12 & 12 & 12 & 12 & 11.04-39.2 & 0.5-27.3 & 1903-1996\\
\addlinespace
SCBI & JUNI & 30 & 30 & 28 & 28 & 20.4-76.19 & 5.6-59.5 & 1903-2010\\
\addlinespace
SCBI & LITU & 106 & 106 & 104 & 104 & 10-91.42 & 0.1-81.1 & 1903-2010\\
\addlinespace
SCBI & PIST & 36 & 36 & 36 & 36 & 13.92-50.96 & 1.6-45.2 & 1931-2010\\
\addlinespace
SCBI & QUAL & 66 & 66 & 66 & 66 & 11.4-76.73 & 0.3-70.4 & 1903-2009\\
\addlinespace
SCBI & QUMO & 67 & 67 & 67 & 67 & 10.22-84.59 & 0.3-69.5 & 1903-2017\\
\addlinespace
SCBI & QURU & 70 & 70 & 70 & 70 & 11.07-87.65 & 2.5-79.2 & 1903-2016\\
\addlinespace
SCBI & QUVE & 81 & 81 & 81 & 81 & 16.02-82.33 & 0.5-78.4 & 1903-2009\\
\addlinespace
SC & PIMA & 442 & 442 & 101 & 101 & 7-14.9 & 0.5-12.5 & 1903-2013\\
\bottomrule
\end{tabular}}
\end{table}
